\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}

\title{Warm Up: Unit analysis and vectors}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Chapter 1 - Unit Analysis and Estimation}

\begin{enumerate}
\item Suppose there are 1000 meters in 1 kilometer, 0.01 meters to one centimeter, and 0.001 meters to 1 millimeter.  Convert the following:
\begin{itemize}
\item 2.3 kilometers to meters
\item 10 meters to centimeters
\item 33 centimeters to meters
\item 0.4 millimeters to meters
\end{itemize}
\item Suppose someone is traveling at a \textit{speed} (ratio of distance to time duration) of 30 km/hr.  Convert this to meters per second. \\ \vspace{1cm}
\end{enumerate}

\section{Chapter 2 - Algebra of Vectors}

\begin{enumerate}
\item A \textit{vector} is a quantity with a \textit{magnitude} and a \textit{direction}.  If we travel 4 km in the x-direction, and we travel 4 km in the y-direction, our displacement from the origin is $\vec{x} = 4 \hat{i} + 4\hat{j}$ km.  Suppose a party of explorers is exploring an area of jungle that we assume to be flat.  The trek is broken into four steps:
\begin{itemize}
\item $\vec{a} = -2 \hat{i} + 2\hat{j}$
\item $\vec{b} = -1 \hat{i} - 1\hat{j}$
\item $\vec{c} = +1 \hat{i} + 5\hat{j}$
\item $\vec{d} = -2 \hat{i} + 0\hat{j}$
\end{itemize}
To obtain the final location of the party, \textit{sum the displacements} $\vec{a}$ through $\vec{d}$.  Where are they? \\ \vspace{1.5cm}
\item Suppose a plane is traveling at 141 km/hr, to the North East.  (a) Draw this vector in a coordinate system below.  (b) How much of the velocity is in the x-direction?  How much is in the y-direction?  \textit{Hint: if East means the x-direction, then the angle with respect to the x-direction is 45 degrees.}
\end{enumerate}

\end{document}
