\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}

\title{Warm Up: Unit analysis and vectors}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Memory Bank}

\begin{enumerate}
\item $\vec{v} = v_x \hat{i} + v_y \hat{j}$ ... Definition of a vector in terms of $\hat{i}$ and $\hat{j}$ components (representing the x-direction and y-direction).
\item $\vec{v} + \vec{w} = (v_x + w_x) \hat{i} + (v_y + w_y) \hat{j}$ ... Vector addition: the $\hat{i}$-components add with each other, and the $\hat{j}$-components add with each other.\
\item $|\vec{v}| = \sqrt{v_x^2 + v_y^2}$ ... The magnitude of the vector
\item $v_x = |\vec{v}| \cos\phi$, $v_y = |\vec{v}| \sin\phi$ ... The x and y-components of the vector
\end{enumerate}

\section{Chapter 1 - Estimation}

\begin{enumerate}
\item Using whatever techniques seem appropriate, \textit{estimate} the following:
\begin{itemize}
\item A: The number of people in class.
\item B: The number of people on this floor of the SLC.
\item C: The number of people in the SLC.
\item D: The number of books in Wardman Library.
\end{itemize}
\end{enumerate}

\section{Chapter 2 - Algebra of Vectors}

\begin{enumerate}
\item Calculate the following:
\begin{itemize}
\item A: If we have a velocity $\vec{v} = -3\hat{i} + 3\hat{j}$ km/hr, what is the \textit{magnitude} of the velocity?
\item B: If we have a velocity $\vec{v} = -3\hat{i} + 3\hat{j}$ km/hr, draw a picture of the velocity in a 2D coordinate system.
\item C: If we have a velocity $\vec{v} = -3\hat{i} + 3\hat{j}$ km/hr, what is the angle the velocity makes with the positive x-axis?
\end{itemize}
\end{enumerate}

\end{document}
