\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}
\usepackage{hyperref}

\begin{document}
\twocolumn

\title{PhET Activity: Work and Energy with the Pendulum}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Memory Bank}

\begin{itemize}
\item Derivative of sine: $\frac{d}{dx} \sin(kx) = k\cos(kx)$
\item Derivative of cosine: $\frac{d}{dx} \cos(kx) = -k\sin(kx)$
\end{itemize}

\section{Introduction}

Let the angle a pendulum makes with the vertical line be $\theta$.  If $\theta \ll 1$, the position of the mass at the end of pendulum is

\begin{align}
x &= L\theta \\
y &= \frac{1}{2}L\theta^2
\end{align}

The gravitational potential energy of the pendulum is 

\begin{align}
U(y) &= mgy \\
U(\theta) &= \frac{1}{2}mgL\theta^2 \\
k &= mgL \\
U(\theta) &= \frac{1}{2}k\theta^2
\end{align}

Notice that the potential energy is a quadratic function, like the potential energy of the spring ($U(x) = \frac{1}{2} k x^2$).  Since $x = L\theta$, $dx = Ld\theta$.  This makes the derivative of $-U$ become

\begin{equation}
F = -\frac{dU}{dx} = -\frac{dU}{Ld\theta} = -mg\theta
\end{equation}

Using Newton's 2nd Law,

\begin{align}
m \frac{d^2 x}{dt^2} &= -mg\theta \\
\frac{d^2 x}{dt^2} &= -g\theta \\
L\frac{d^2 \theta}{dt^2} &= -g\theta \\
\frac{d^2 \theta}{dt^2} &= -\left(\frac{g}{L}\right)\theta \label{eq:1}
\end{align}

Let the \textit{angular frequency} $\omega$ be defined by $\omega^2 = g/L$.  Equation \ref{eq:1} becomes

\begin{equation}
\frac{d^2 \theta}{dt^2} = -\omega^2\theta \label{eq:2}
\end{equation}

\textbf{Equation \ref{eq:2} produces a sum of sines and cosines.}

\section{Physical Pendulum Behavior}

Notice the form of the physical pendulum constructed in the lab.  A ruler is suspended with a string, clamps, and rods.  The ruler has a mass of 87 grams.  We can tape the 50 gram weight to various locations along the ruler.

\begin{enumerate}
\item Create a graph below of the \textit{period} of the pendulum versus the location of the 50 gram weight. \\ \vspace{5cm}
\item Save a table of the \textit{period} of the pendulum versus the location of the 50 gram weight below. \\ \vspace{5cm}
\item How can you explain the trend in the data using torque, moment of inertia, and Eq. \ref{eq:2}?
\end{enumerate}

\end{document}
