\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}
\twocolumn

\title{Monday Warm Up: Unit 6: Rotational Motion I}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Memory Bank}

\begin{itemize}
\item $s = r \theta$ ... Definition of the radian.
\item $\vec{s} = \vec{\theta} \times \vec{r}$ ... Vector relationship between angular displacement, arc length, and radius.
\item $v = r \omega$ ... Relationship between tangential velocity, angular velocity, and radius.
\item $\vec{v} = \vec{\omega} \times \vec{r}$ ... Vector relationship between tangential velocity, angular velocity, and radius.
\end{itemize}

\section{Rotational Variables and Rotational Motion}

\begin{enumerate}
\item Suppose the displacement of an aircraft along a curved trajectory is $\vec{s} = 0.1 \hat{j}$ km, and the radius from our origin is $\vec{r} = 10 \hat{i}$ km. What is $\vec{\theta}$? \\ \vspace{1cm}
\item Suppose the displacement of an aircraft along a curved trajectory is $\vec{s} = -0.1 \hat{i}$ km, and the radius from our origin is $\vec{r} = 1 \hat{j}$ km. What is $\vec{\theta}$? \\ \vspace{1cm}
\item Suppose the angular velocity of our aircraft is $\vec{\omega} = + \pi/2 \hat{k}$ radians/minute, and the radius of curvature of our trajectory is $\vec{r} = -10\hat{i}$ km.  What is our velocity vector? \\ \vspace{1cm}
\item Perform the following unit conversions:
\begin{itemize}
\item 45 rpm to radians per second.
\item 100 radians per second to rpm.
\item 6000 rpm to rotations per millisecond.
\end{itemize} \vspace{2cm}
\item A particle moves 3.0 m along a circle of radius 1.5 m. (a) Through what angle does it rotate? (b) If the particle makes this trip in 1.0 s at a constant speed, what is its angular velocity? (c) What is its acceleration? \\ \vspace{2cm} 
\item  On takeoff, the propellers on a UAV (unmanned aerial vehicle) increase their angular velocity for 3.0 s from rest at a rate of $\omega=(25.0t)$ rad/s where $t$ is measured in seconds. (a) What is the instantaneous angular velocity of the propellers at t=2.0s? (b) What is the angular acceleration?
\end{enumerate}

\end{document}
