\title{Midterm 4 for Calculus-Based Physics-1: Mechanics (PHYS150-01)}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{November 27th, 2017}
\documentclass[10pt]{article}
\usepackage[a4paper, total={18cm, 27cm}]{geometry}
\usepackage{outlines}
\usepackage[sfdefault]{FiraSans}
\usepackage{graphicx}

\begin{document}
\maketitle

\section{Definition of Momentum}
\begin{enumerate}
\item Which of the following quantities has units of momentum?
\begin{itemize}
\item 4 m/s
\item 8 kg m/s
\item 10 m/s$^2$
\item 3 N m
\end{itemize}
\item Which two properties of a system must be true for momentum to be conserved?
\begin{itemize}
\item The mass must be constant, and there must be no net external force.
\item The mass must be constant, and there must be no internal forces.
\item There must be no net external force, and the velocity vectors must sum to zero.
\item There must be no internal forces and no net external force.
\end{itemize}
\item In which of the following situations will momentum be conserved?
\begin{itemize}
\item An object skids to a stop due to surface friction.
\item After a rocket has launched vertically upward, and has run out of fuel.
\item A hockey puck glides along frictionless ice and rebounds off of a wall.
\item A sheet of paper falls to the ground at constant velocity and then rests on the ground.
\end{itemize}
\item What is the momentum of a 10 kg meteorite traveling through space at 1 km/s? \\ \vspace{1cm}
\end{enumerate}
\section{Conservation of Momentum}
\begin{enumerate}
\item A 100 gram meteorite breaks into two smaller ones.  The two final velocities are observed to be equal, and in the same direction as the original velocity.  What are the masses of the two pieces, if momentum is conserved? \\ \vspace{1.5cm}
\item Two protons interact and scatter off of each other.  The mass of both particles cannot change.  If one proton has a velocity of $v_{\rm 1} = -10^{6}$ m/s, and the other has $v_{\rm 2} = 10^{6}$ m/s, what are the final velocities?  Draw a diagram of the initial state and the final state.  \\ \vspace{1.5 cm}
\end{enumerate}
\section{Classifying Interactions}
\begin{enumerate}
\item What quantity is conserved in an \textit{elastic} interaction but not in an \textit{inelastic} interaction?
\begin{itemize}
\item Gravitational potential energy
\item Kinetic energy
\item Momentum
\item Internal forces
\end{itemize}
\item If an interaction is \textit{totally} inelastic, what is the final velocity?
\begin{itemize}
\item Less than the initial velocity
\item Such that the kinetic energy decreases
\item Such that the final momentum is zero
\item The final velocity is zero
\end{itemize}
\item The notation $n \rightarrow n$ means many-to-many, or several particles scattering off of each other.  Which of the following is true of $n \rightarrow n$ interactions?
\begin{itemize}
\item They must be at least partially elastic
\item They can only be totally elastic
\item They must be inelastic
\item They must be totally inelastic
\end{itemize}
\item The notation $n \rightarrow n$ means many-to-many, or several particles scattering off of each other.  Which of the following is true of $n \rightarrow n$ interactions?
\begin{itemize}
\item They must be elastic
\item They cannot be inelastic
\item They must be inelastic
\item They cannot be inelastic
\end{itemize}
\item A 1 kg particle has $v_{\rm 1} = 0$ m/s, and it interacts with a 2 kg particle with velocity $v_{\rm 2} = 2$ m/s.  If the collision is inelastic, what is the final velocity of the combined system? \\ \vspace{1.5cm}
\item A 1 kg particle has $v_{\rm 1} = -1$ m/s, and it interacts with a 1 kg particle with velocity $v_{\rm 2} = 1$ m/s.  If the collision is elastic, what is the final velocity of each particle? \\ \vspace{1.5 cm}
\end{enumerate}
\end{document}