\documentclass{article}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{amsmath}

\begin{document}

\title{Monday Reading Assessment: Unit 5, Field Induction and Inductance}
\author{Prof. Jordan C. Hanson}

\maketitle

\section{Memory Bank}

\begin{itemize}
\item $\epsilon = -N d\phi_m /dt$ ... Faraday's Law
\item $\phi_m = \vec{B} \cdot \vec{A} = BA \cos(\theta)$ ... Definition of magnetic flux
\item $\epsilon(t) = V_{max} \sin(\omega t)$ ... AC voltage signal.
\item $\omega = 2\pi f$ ... Angular frequency vs. frequency.
\end{itemize}

\section{Electric guitar pickup coils}

\begin{enumerate}
\item Consider a \textit{pickup coil} for an electric guitar that has $N = 1000$ turns of wire around a permanent magnet with $\vec{B} = 0.03\hat{k}$ T. The area of the pickup coil is $\vec{A} = 10^{-4}\hat{k}$ cm$^{-2}$. The permanent magnet \textit{magnetizes} the guitar string above it (gives it a magnetic dipole moment).  If the string oscillates with a frequency $\omega$, then we can model the B-field of string as
\begin{equation}
\vec{b} = -b_0\sin(\omega t)\hat{k}
\end{equation}
(a) If $b_0 = 0.01$ T, and $\omega = 1350$ Hz, what is the total magnetic field function?  (b) Graph the magnetic field as a function of time. \\ \vspace{2cm}
\item What is the induced voltage in the pickup coil by the metal string as it vibrates? (Use the given variables from the previous problem). \\ \vspace{2cm}
\item What is the frequency $f$ of the pickup coil voltage? \\ \vspace{1cm}
\item From the pickup coil, the induced voltage is passed to an analog \textit{amplifier}, which grows the signal by a factor of 200.  What is the $V_{max}$ of the signal after it has been amplified by a factor of 200?
\end{enumerate}
\end{document}
